Viscous Erosion in Porous Media

### 2018-03-11 docs/methods/erosion_methods.tex
small changes to BIE section

A challenge with erosion is that it creates corners at points on the
body where the shear stress vanishes.  There is recent work for solving
boundary integral equations in geometries with corners by using
carefully chosen local expansions around each corner~\cite{rac-ser2017,
ser-rok2016} or panel-based graded meshes~\cite{hel2011,
gil-hao-mar2014, bre2012}.  The limitation of these methods is that they
require the corner's location, a non-uniform spaced mesh, and some of
the methods have a large condition number.  So that we can use a uniform
grid, we instead slightly modify the geometry to maintain sufficient
regularity.  First, we add a regularization term to the shear stress to
penalize regions of high curvature.  Second, we apply a Gaussian filter
to smooth the absolute value of the shear stress before the body shape
is updated.

@@

%\todo[inline]{These are some of my thoughts on reworking this first paragraph. See if you like it.}
We begin by making two fundamental design choices: first, we restrict ourselves to two spatial dimensions, and second the bodies are considered immobile (i.e.~fixed to their initial locations). The restriction to 2D allows us to simulate the erosion of many bodies (order 50-100) with high resolution, giving us the ability to accurately characterize the erosion of a porous medium. While BIE methods are available in 3D, the extra computational cost limits the number of bodies that could be practically simulated, making porous-medium simulations less feasible. The second assumption of immobile bodies affords a simple framework to study the effects of erosion in isolation. While other effects, such as sedimentation or transport, undoubtedly play a role in real applications, these effects would introduce additional complexity, both conceptually and numerically.

@@

In this section, we describe the boundary integral equations employed
for the fluid equations \cite{Moore2007}
\todo[inline]{Bryan do you want to add a bunch of references to your boundary integral papers here (or somewhere else)?}
In particular, we solve the incompressible Stokes equations
with Dirichlet boundary conditions 
\begin{alignat*}{3}
  \mu\Delta \uu &= \grad p, \qquad && \xx \in \Omega, \\
  \grad \cdot \uu &= 0,   && \xx \in \Omega, \\
  \uu &= \ff,  && \xx \in \bd\Omega.
\end{alignat*}
The boundary condition $\ff$ is the prescribed velocity $\UU$ on
$\Gamma$, and it is zero on the grains $\gamma$.  As the viscosity $\mu$
also sets the time scale, we take $\mu=1$.  We start by defining the
double-layer potential
\todo[inline]{Do you think we should flesh out this time-scale stuff more precisely? i.e.~By setting $\mu = 1$ and $C = 1$ (the erosion constant), what exactly is the timescale that defines our dimensionless time.}
\begin{align}
  \DD[\eeta](\xx) = \frac{1}{\pi} \int_{\bd\Omega} 
    \frac{\rr \cdot \nn}{\rho^2} \frac{\rr \otimes \rr}{\rho^2} 
    \eeta(\yy) \, ds_\yy, \quad \xx \in \Omega,
    \label{eqn:velocityDLP}
\end{align}
where $\rr = \xx - \yy$, $\rho = \|\rr\|$, and $\eeta$ is an unknown density function.  The
double-layer potential cannot capture velocity fields corresponding to
rigid body motions~\cite{pow-mir1987}.  The rank deficiency is removed
by complementing the double-layer potential with a Stokeslets and
rotlets centered inside each interior body
\begin{align}
  S[\llambda_\ell,\cc_\ell] = \frac{1}{4\pi} \left(-\log \rho \II + 
    \frac{\rr \otimes \rr}{\rho^2} \right)\llambda_\ell
  \quad \text{and} \quad
  R[\xi_\ell,\cc_\ell] = \frac{\rr^\perp}{\rho^2}\xi_\ell, 
  \quad \ell=1,\ldots,M,
  \label{eqn:stokeslet_rotlet}
\end{align}
where  $\rr = \xx - \cc_\ell$, $\cc_\ell$ is a point inside the
$\ell^{th}$ body, $\rr^\perp = (r_2,-r_1)$, and $\rho = \|\rr\|$.  Then,
any solution of the Stokes equations with a Dirichlet boundary condition
$\ff$ can be written in terms of the completed double-layer potential
\begin{align}
  \label{eqn:completed_DLP}
  \uu(\xx) = \DD[\eeta](\xx) + 
    \sum_{\ell=1}^{M} S[\llambda_\ell,\cc_\ell](\xx) +
    \sum_{\ell=1}^{M} R[\xi_\ell,\cc_\ell](\xx), \quad \xx \in \Omega,
\end{align}
where the density function, Stokeslets, and rotlets satisfy
\begin{subequations}
\label{eqn:completed_BIE}
\begin{align}
  \ff(\xx) &= -\frac{1}{2}\eeta(\xx) + \DD[\eeta](\xx) + 
      \sum_{\ell=1}^{M} S[\llambda_\ell,\cc_\ell](\xx) +
      \sum_{\ell=1}^{M} R[\xi_\ell,\cc_\ell](\xx) + \NN_0[\eeta](\xx),
      \quad \xx \in \bd\Omega, 
      \label{eqn:DLP}\\
  \llambda_\ell &= \frac{1}{2\pi}\int_{\gamma_\ell} \eeta(\yy) \, ds_\yy,
  \quad \ell=1,\ldots,M,
  \label{eqn:stokeslet} \\
  \xi_\ell &= \frac{1}{2\pi}\int_{\gamma_\ell} \yy^\perp \cdot \eeta(\yy)
  \, ds_\yy, \quad \ell=1,\ldots,M.
  \label{eqn:rotlet}
\end{align}
\end{subequations}
Here,
\begin{align*}
  \NN_0[\eeta](\xx) = \int_{\Gamma} 
    (\nn(\xx) \otimes \nn(\yy))\eeta(\yy) \, ds_\yy
\end{align*}
removes the rank one null space that results from the compatibility
constraint associated with the flux-free requirement of $\ff$.

>>>>

A challenge with erosion is that it creates corners at points on the
body where the shear stress vanishes.  There is recent work for solving
boundary integral equations in geometries with corners by using
carefully chosen local expansions around each corner~\cite{rac-ser2017,
ser-rok2016} or panel-based graded meshes~\cite{hel2011,
gil-hao-mar2014, bre2012}.  The limitation of these methods is that they
require the corner's location, a non-uniform spaced mesh, and some of
the methods have a large condition number.  One solution is to use two
meshes---one for the quadrature the other for the interface tracking.
Then, interpolation between the two grids is required, and this is the
approach taken by Ojala and Tornberg~\cite{oja-tor2015}.  Instead of
using two grids and interpolation operators, we only require the uniform
grid by slightly modify the geometry to maintain sufficient regularity.
First, we add a regularization term to the shear stress to penalize
regions of high curvature.  Second, we apply a Gaussian filter to smooth
the absolute value of the shear stress before the body shape is updated.

@@

%\todo[inline]{These are some of my thoughts on reworking this first paragraph. See if you like it.}
We begin by making two fundamental design choices: first, we restrict
ourselves to two spatial dimensions, and second the bodies are
considered immobile (i.e.~fixed to their initial locations). The
restriction to 2D allows us to simulate the erosion of many bodies
(up to $50$ in this manuscript) with high resolution, giving us the
ability to accurately characterize the erosion of a porous medium. While
boundary integral equation (BIE) methods are available in 3D, the extra computational cost limits the number of bodies that could be practically simulated, making porous-medium simulations less feasible. The second assumption of immobile bodies affords a simple framework to study the effects of erosion in isolation. While other effects, such as sedimentation or transport, undoubtedly play a role in real applications, these effects would introduce additional complexity, both conceptually and numerically.

@@

In this section, we describe the boundary integral equations employed
for the fluid equations.  We will start by writing the velocity field in
terms of the standard double-layer potential~\cite{laz1963,poz1992}.
Since the double-layer potential representation is only able to
represent force-free and torque-free velocity fields, Power and
Miranda~\cite{pow-mir1987} completed the BIE formulation by introducing
a Stokeslet and Rotlet.  This methodology was extended by Power for
multiply-connected domains~\cite{pow1993}.  In two-dimensions, a stream
function formulation can be used in which case the biharmonic equation
must be solved, and this can be done using BIEs by using analytic
function theory~\cite{gre-kro-may1996}.

  \cite{Moore2007}
\todo[inline]{Bryan do you want to add a bunch of references to your boundary integral papers here (or somewhere else)?}
In particular, we solve the incompressible Stokes equations
with Dirichlet boundary conditions 
\begin{alignat*}{3}
  \mu\Delta \uu &= \grad p, \qquad && \xx \in \Omega, \\
  \grad \cdot \uu &= 0,   && \xx \in \Omega, \\
  \uu &= \ff,  && \xx \in \bd\Omega.
\end{alignat*}
The boundary condition $\ff$ is the prescribed velocity $\UU$ on
$\Gamma$, and it is zero on the grains $\gamma$.  As the viscosity $\mu$
also sets the time scale, we take $\mu=1$.  We start by defining the
double-layer potential
\todo[inline]{Do you think we should flesh out this time-scale stuff more precisely? i.e.~By setting $\mu = 1$ and $C = 1$ (the erosion constant), what exactly is the timescale that defines our dimensionless time.}
\begin{align}
  \DD[\eeta](\xx) = \frac{1}{\pi} \int_{\bd\Omega} 
    \frac{\rr \cdot \nn}{\rho^2} \frac{\rr \otimes \rr}{\rho^2} 
    \eeta(\yy) \, ds_\yy, \quad \xx \in \Omega,
    \label{eqn:velocityDLP}
\end{align}
where $\rr = \xx - \yy$, $\rho = \|\rr\|$, and $\eeta$ is an unknown density function.  The
double-layer potential cannot capture velocity fields corresponding to
rigid body motions~\cite{pow-mir1987}.  The rank deficiency is removed
by complementing the double-layer potential with a Stokeslets and
rotlets centered inside each interior body
\begin{align}
  S[\llambda_\ell,\cc_\ell] = \frac{1}{4\pi} \left(-\log \rho \II + 
    \frac{\rr \otimes \rr}{\rho^2} \right)\llambda_\ell
  \quad \text{and} \quad
  R[\xi_\ell,\cc_\ell] = \frac{\rr^\perp}{\rho^2}\xi_\ell, 
  \quad \ell=1,\ldots,M,
  \label{eqn:stokeslet_rotlet}
\end{align}
where  $\rr = \xx - \cc_\ell$, $\cc_\ell$ is a point inside the
$\ell^{th}$ body, $\rr^\perp = (r_2,-r_1)$, and $\rho = \|\rr\|$.  Then,
any solution of the Stokes equations with a Dirichlet boundary condition
$\ff$ can be written in terms of the completed double-layer potential
\begin{align}
  \label{eqn:completed_DLP}
  \uu(\xx) = \DD[\eeta](\xx) + 
    \sum_{\ell=1}^{M} S[\llambda_\ell,\cc_\ell](\xx) +
    \sum_{\ell=1}^{M} R[\xi_\ell,\cc_\ell](\xx), \quad \xx \in \Omega,
\end{align}
where the density function, Stokeslets, and rotlets satisfy
\begin{subequations}
\label{eqn:completed_BIE}
\begin{align}
  \ff(\xx) &= -\frac{1}{2}\eeta(\xx) + \DD[\eeta](\xx) + 
      \sum_{\ell=1}^{M} S[\llambda_\ell,\cc_\ell](\xx) +
      \sum_{\ell=1}^{M} R[\xi_\ell,\cc_\ell](\xx) + \NN_0[\eeta](\xx),
      \quad \xx \in \bd\Omega, 
      \label{eqn:DLP}\\
  \llambda_\ell &= \frac{1}{2\pi}\int_{\gamma_\ell} \eeta(\yy) \, ds_\yy,
  \quad \ell=1,\ldots,M,
  \label{eqn:stokeslet} \\
  \xi_\ell &= \frac{1}{2\pi}\int_{\gamma_\ell} \yy^\perp \cdot \eeta(\yy)
  \, ds_\yy, \quad \ell=1,\ldots,M.
  \label{eqn:rotlet}
\end{align}
\end{subequations}
Here,
\begin{align*}
  \NN_0[\eeta](\xx) = \int_{\Gamma} 
    (\nn(\xx) \otimes \nn(\yy))\eeta(\yy) \, ds_\yy
\end{align*}
removes the rank one null space that results from the compatibility
constraint associated with the flux-free requirement of $\ff$.
